Software for the analysis and visualization of deep mutational scanning data

### 2015-03-24 paper/BMC_Bioinformatics_2015/bmc_bioinformatics_article_nofigs.tex

\section*{Background}
Deep mutational scanning is a high-throughput experimental technique to assess the impacts of mutations on a protein-coding gene~\cite{fowler2014deep}. Figure \ref{fig:DeepMutationalScan} shows a schematic of deep mutational scanning. A gene is mutagenized, and the library of resulting variants is introduced into cells or viruses, which are then subjected to an experimental selection that enriches for functional variants and depletes non-functional ones. Deep sequencing of the variants pre- and post-selection provides information about the functional impacts of mutations. Since the original description of deep mutational scanning by Fowler \textit{et al}~\cite{fowler2010high}, the technique has been applied to a wide range of genes~\cite{traxlmayr2012construction,mclaughlin2012spatial,starita2013activity,melamed2013deep,roscoe2013analyses,firnberg2014comprehensive,bloom2014experimentally,melnikov2014comprehensive,thyagarajan2014inherent,wu2014high,wu2014ns,olson2014comprehensive}, both to measure mutational tolerance given a single selection pressure as in Figure \ref{fig:DeepMutationalScan}A, or to identify mutations that have different effects under alternative selections as in Figure \ref{fig:DeepMutationalScan}B. New techniques to create comprehensive codon-mutant libraries of genes make it possible to profile all single amino-acid mutations~\cite{firnberg2012,jain2013rapid,firnberg2014comprehensive,bloom2014experimentally,melnikov2014comprehensive}, while new techniques for targeted mutagenesis of mammalian genomes enable deep mutational scanning to be applied across the biological spectrum from viruses and bacteria to human cells~\cite{findlay2014saturation}.

@@

The data for each site $r$ is characterized by the sequencing \emph{depth} (total number of counts); let $N_r^{\rm{pre}}$, $N_r^{\rm{post}}$, $N_r^{s1}$, and $N_r^{s2}$ denote the depth at $r$ for each of the four libraries in Figure \ref{fig:DeepMutationalScan} (pre-selection, post-selection, selection $s1$, and selection $s2$). Typical depths for current experiments are $N \sim 10^6$. Denote the counts of character $x$ (characters might be nucleotides, amino acids, or codons) at $r$ as $n_{r,x}^{\rm{pre}}$, $n_{r,x}^{\rm{post}}$, $n_{r,x}^{s1}$, and $n_{r,x}^{s2}$. The values of $n_{r,x}$ for characters $x$ that differ from the wildtype identity $\operatorname{wt}\left(r\right)$ depend on both the depth $N$ and the average per-site mutation rate. Typical experiments introduce about one mutation per gene, so the average mutation rate is $\overline{\mu} \sim 1/L$ where $L$ is the length of the gene. Therefore, if a 500-codon gene is sequenced at depth $N \sim 10^6$, we expect $N \overline{\mu} \sim 2000$ counts of non-wildtype codons at each site. Since there are 63 mutant codons, the average pre-selection counts for a mutation to a specific $x \ne \operatorname{wt}\left(r\right)$ will be $n_{r,x}^{\rm{pre}} \sim 30$, with counts for most mutations deviating from this average due to biases in creation of the mutant library and randomness in which molecules are sequenced. Counts in the post-selection libraries will further deviate from this average due to selection. Therefore, even at depths $N \sim 10^6$, the actual counts of most mutations will be quite modest.

\subsection*{The goal: inferring site-specific amino-acid preferences.}
The goal is to estimate the effects of mutations from changes in their counts after selection. Let $\mu_{r,x}$, $f_{r,x}$, $f_{r,x}^{s1}$, and $f_{r,x}^{s2}$ denote the \emph{true} frequencies of $x$ at $r$ that would be observed for the four libraries in Figure \ref{fig:DeepMutationalScan} if we sampled at infinite depth in both the actual experiment and the sequencing. If we knew these frequencies, we could calculate parameters that reflect the effects of mutations. One parameter that characterizes the effect of mutating $r$ from $\operatorname{wt}\left(r\right)$ to $x$ for the experiment in Figure \ref{fig:DeepMutationalScan}A is the \emph{enrichment ratio}, which is the relative frequency of mutations to $x$ after selection versus before selection:
\begin{equation}
\label{eq:enrichment}
\phi_{r,x} = \frac{f_{r,x} / f_{r,\operatorname{wt}\left(r\right)}}{\mu_{r,x} / \mu_{r,\operatorname{wt}\left(r\right)}}.
\end{equation}
Beneficial mutations have $\phi_{r,x} > 1$, while deleterious ones have $\phi_{r,x} < 1$. A related parameter is the \emph{preference} $\pi_{r,x}$ of $r$ for $x$. At each site, the preferences are simply the enrichment ratios rescaled to sum to one:
\begin{equation}
\label{eq:preference}
\pi_{r,x} = \frac{\phi_{r,x}}{\sum_y \phi_{r,y}} = \frac{f_{r,x} / \mu_{r,x}}{\sum_y f_{r,y} / \mu_{r,y}},
\end{equation}
or equivalently
\begin{equation}
\label{eq:preference2}
f_{r,x} = \frac{\pi_{r,x} \times \mu_{r,x}}{\sum_y \pi_{r,y} \times \mu_{r,y}},
\end{equation}
where $y$ is summed over all character identities (all nucleotides, codons, or amino acids).
The preferences can be intuitively visualized (Figure \ref{fig:DeepMutationalScan}A) and interpreted as the equilibrium frequencies in substitution models for gene evolution~\cite{bloom2014experimentally,bloom2014informed} (after accounting for uneven mutational rates~\cite{yampolsky2005exchangeability,stoltzfus2009climbing}).

@@

\subsubsection*{Implementation.}
The program \texttt{dms\_inferprefs} in the \texttt{dms\_tools} package infers the preferences by using \texttt{pystan}~\cite{pystan-software:2014} to perform MCMC over the posterior defined by the product of the likelihoods and priors in Equations \ref{eq:pr_nrerrpre}, \ref{eq:pr_nrerrpost}, \ref{eq:pr_nrpre}, \ref{eq:pr_nrpost}, \ref{eq:pr_pir}, \ref{eq:pr_mur}, \ref{eq:pr_epsilonr}, and \ref{eq:pr_rhor}. The program runs four chains from different initial values, and checks for convergence by ensuring that the Gelman-Rubin statistic $\hat{R}$~\cite{gelman1992inference} is $<1.1$ and the effective sample size is $>100$; the number of MCMC iterations is increased until these convergence is achieved. The program \texttt{dms\_logoplot} in the \texttt{dms\_tools} package visualizes the posterior mean preferences via an extension to \texttt{weblogo}~\cite{crooks2004}.

@@

Figures \ref{fig:MelnikovPrefs} and \ref{fig:ThyagarajanPrefs} also illustrate use of \texttt{dms\_correlate} to assess the correlation between preferences inferred from different biological replicates~\cite{blainey2014points} of the experiment. The inclusion and analysis of such replicates provide the only sure way to fully assess the sources of noise associated with deep mutational scanning.

\subsubsection*{Testing on simulated data.}
To test the accuracy of preference-inference by \texttt{dms\_tools}, I simulated deep mutational scanning counts using the preferences in Figure \ref{fig:MelnikovPrefs}, both with and without errors quantified by appropriate controls. I then used \texttt{dms\_tools} to infer preferences from the simulated data, and also made similar inferences using simple ratio estimation (Equation \ref{eq:enrichmentcounts}). Figure \ref{fig:PrefCorrelations} shows the inferred preferences versus the actual values used to simulate the data. For simulations with mutation counts (quantified by the product $N\overline{\mu}$ of the depth and average per-site mutation rate) $\sim 1000$ to 2000, the inferences are quite accurate. Inferences made by \texttt{dms\_tools} are always more accurate than those obtained by simply taking ratios of mutation counts.

@@

\subsubsection*{Implementation.}
The program \texttt{dms\_inferdiffprefs} in the \texttt{dms\_tools} package infers the differential preferences by performing MCMC over the posterior defined by the product of the likelihoods and priors in Equations \ref{eq:pr_nrerr}, \ref{eq:pr_nrstart}, \ref{eq:pr_nrs1}, \ref{eq:pr_nrs2}, \ref{eq:pr_controlpir}, \ref{eq:pr_xir}, \ref{eq:pr_frstart}, and \ref{eq:pr_deltapi}. The MCMC is performed as described for the preferences, and characters can again be any of nucleotides, amino acids, or codons. The program \texttt{dms\_logoplot} visualizes the posterior mean differential preferences via an extension to \texttt{weblogo}~\cite{crooks2004}. In addition, \texttt{dms\_inferdiffprefs} also creates text files that give the posterior probability that $\Delta\pi_{r,x} > 0$ or $< 0$ (in other words, the probability that there is differential selection).

\subsection*{Inferring differential preference with \texttt{dms\_tools}.}
To test the accuracy of differential preference inference by \texttt{dms\_tools}, I simulated an experiment like that in Figure \ref{fig:DeepMutationalScan}B with the starting counts based on Melnikov \textit{et al}'s actual deep mutational scanning data of a Tn5 transposon~\cite{melnikov2014comprehensive}. As show by Figure \ref{fig:MelnikovSimulatedDiffPrefs}, \texttt{dms\_inferdiffprefs} accurately infers the differential preferences at typical experimental depths. The results are easily visualized with \texttt{dms\_logoplot}. To provide a second illustration of differential preferences, Figure \ref{fig:WuDiffPrefs} shows an analysis of the data obtained by Wu \textit{et al} when they performed an experiment like that in Figure \ref{fig:DeepMutationalScan}B on nucleotide mutants of the influenza NS gene in the presence or absence of interferon treatment.

@@

\begin{figure}[h!]
%\centerline{\includegraphics[width=6.5in]{../../examples/Melnikov_et_al_Tn5/infer_prefs_on_simulated_data/correlations.pdf}}
\cprotect\caption{\label{fig:PrefCorrelations}\csentence{Accuracy of preference inference on simulated data.} Deep mutational scanning counts were simulated using the preferences in Figure \ref{fig:MelnikovPrefs}A and realistic parameters for mutation and error rates. The simulations were done {\bf (A)} without or {\bf (B)} with sequencing errors quantified by control libraries. Plots show the correlation between the actual and inferred preferences as a function of the product of the sequencing depth $N$ and the average per-site mutation rate $\overline{\mu}$; real experiments typically have $N\overline{\mu} \sim 1000$ to 2000 depending on the sequencing depth and gene length. Preferences are inferred using the full algorithm in \texttt{dms\_tools} (top panels) or by simply calculating ratios of counts (bottom panels) using Equation \ref{eq:enrichmentcounts} and its logical extension to include errors, both with a pseudocount of one. The \texttt{dms\_tools} inferences are more accurate than the simple ratio estimation, with both methods converging to the actual values with increasing $N\overline{\mu}$. Given files with the mutation counts, the plots in this figure can be generated as \texttt{prefs\_corr.pdf} and \texttt{ratio\_corr.pdf} with commands such as:
\begin{lstlisting}
dms_inferprefs pre.txt post.txt inferred_prefs.txt --ncpus -1
dms_inferprefs pre.text post.text ratio_prefs.txt --ratio_estimation 1
dms_correlate actual_prefs.txt inferred_prefs.txt prefs_corr --name1 "actual" --name2 "inferred" --corr_on_plot --r2
dms_correlate actual_prefs.txt ratio_prefs.txt ratio_corr --name1 "actual" --name2 "inferred" --corr_on_plot --r2
\end{lstlisting}
}
\end{figure}

>>>>

\section*{Background}
Deep mutational scanning is a high-throughput experimental technique to assess the impacts of mutations on a protein-coding gene~\cite{fowler2014deep}. Figure \ref{fig:DeepMutationalScan} shows a schematic of deep mutational scanning. A gene is mutagenized, and the library of resulting variants is introduced into cells or viruses, which are then subjected to an experimental selection that enriches for functional variants and depletes non-functional ones. Deep sequencing of the variants pre- and post-selection provides information about the functional impacts of mutations. Since the original description of deep mutational scanning by Fowler \textit{et al}~\cite{fowler2010high}, the technique has been applied to a wide range of genes~\cite{traxlmayr2012construction,mclaughlin2012spatial,starita2013activity,melamed2013deep,roscoe2013analyses,firnberg2014comprehensive,bloom2014experimentally,melnikov2014comprehensive,thyagarajan2014inherent,wu2014high,wu2014ns,olson2014comprehensive,kitzman2015massively}, both to measure mutational tolerance given a single selection pressure as in Figure \ref{fig:DeepMutationalScan}A, or to identify mutations that have different effects under alternative selections as in Figure \ref{fig:DeepMutationalScan}B. New techniques to create comprehensive codon-mutant libraries of genes make it possible to profile all amino-acid mutations~\cite{firnberg2012,jain2013rapid,firnberg2014comprehensive,bloom2014experimentally,melnikov2014comprehensive,kitzman2015massively}, while new techniques for targeted mutagenesis of mammalian genomes enable deep mutational scanning to be applied across the biological spectrum from viruses and bacteria to human cells~\cite{findlay2014saturation}.

@@

The data for each site $r$ is characterized by the sequencing \emph{depth} (total number of counts); let $N_r^{\rm{pre}}$, $N_r^{\rm{post}}$, $N_r^{s1}$, and $N_r^{s2}$ denote the depth at $r$ for each of the four libraries in Figure \ref{fig:DeepMutationalScan} (pre-selection, post-selection, selection $s1$, and selection $s2$). Typical depths for current experiments are $N \sim 10^6$. Denote the counts of character $x$ (characters might be nucleotides, amino acids, or codons) at $r$ as $n_{r,x}^{\rm{pre}}$, $n_{r,x}^{\rm{post}}$, $n_{r,x}^{s1}$, and $n_{r,x}^{s2}$. The values of $n_{r,x}$ for characters $x$ that differ from the wildtype identity $\operatorname{wt}\left(r\right)$ depend on both the depth $N$ and the average per-site mutation rate $\overline{\mu}$. Since the mutations are intentionally introduced into the mutant library by the experimentalist, in principle $\overline{\mu}$ could have any value. But typically, deep mutational scanning experiments aim to introduce about one mutation per gene to avoid filling the mutant library with highly mutated genes -- so the average mutation rate is usually $\overline{\mu} \sim 1/L$ where $L$ is the length of the gene. Therefore, if a 500-codon gene is sequenced at depth $N \sim 10^6$, we expect $N \overline{\mu} \sim 2000$ counts of non-wildtype codons at each site. Since there are 63 mutant codons, the average pre-selection counts for a mutation to a specific $x \ne \operatorname{wt}\left(r\right)$ will be $n_{r,x}^{\rm{pre}} \sim 30$, with counts for most mutations deviating from this average due to biases in creation of the mutant library and randomness in which molecules are sequenced. Counts in the post-selection libraries will further deviate from this average due to selection. Therefore, even at depths $N \sim 10^6$, the actual counts of most mutations will be quite modest.

The rest of this paper assumes that the sequencing depth is less than the number of unique molecules in the mutant library, such that the deep sequencing randomly subsamples the set of molecules. If this assumption is false (i.e. if the number of unique molecules is substantially less than the sequencing depth), then the accuracy of inferences about mutational effects will be fundamentally limited by this aspect of the experimental design. Properly done experiments should quantify the number of unique molecules in the library so that it is obvious whether this assumption holds. In the absence of such information, the analysis can be repeated using only a random fraction of the deep sequencing data to assess whether inferences are limited by sequencing depth or the underlying molecular diversity in the mutant library.

\subsection*{The goal: inferring site-specific amino-acid preferences.}
The goal is to estimate the effects of mutations from changes in their counts after selection. Let $\mu_{r,x}$, $f_{r,x}$, $f_{r,x}^{s1}$, and $f_{r,x}^{s2}$ denote the \emph{true} frequencies at site $r$ of all mutant characters $x \ne \operatorname{wt}\left(r\right)$ that would be observed for the four libraries in Figure \ref{fig:DeepMutationalScan} if we sampled at infinite depth in both the actual experiment and the sequencing. The definition of these frequencies for the wildtype character $\operatorname{wt}\left(r\right)$ depends on how the mutant library is constructed. If the mutant library is constructed so that there is a Poisson distribution of the number of mutations per gene (as is the case for error-prone PCR or the codon-mutagenesis in \cite{bloom2014experimentally,thyagarajan2014inherent}), then $\mu_{r,\operatorname{wt}\left(r\right)}$, $f_{r,\operatorname{wt}\left(r\right)}$, $f_{r,\operatorname{wt}\left(r\right)}^{s1}$, and $f_{r,\operatorname{wt}\left(r\right)}^{s2}$ are defined as for all other characters $x$, and denote the frequencies of $\operatorname{wt}\left(r\right)$ at site $r$ that would be observed if sampling at infinite depth. The reason we can make this definition for libraries containing genes with Poisson-distributed numbers of mutations is that for any reasonable-length gene ($L \gg 1$), the marginal distribution of the number of mutations in a gene is virtually unchanged by the knowledge that there is a mutation at site $r$. On the other hand, if the mutant library is constructed so that there is exactly zero or one mutation per gene (as in \cite{firnberg2014comprehensive,melnikov2014comprehensive,kitzman2015massively}), then the marginal distribution of the total number of mutations in a gene is changed by the knowledge that there is a mutation at $r$. In this case, the wildtype-character frequencies $\mu_{r,\operatorname{wt}\left(r\right)}$, $f_{r,\operatorname{wt}\left(r\right)}$, $f_{r,\operatorname{wt}\left(r\right)}^{s1}$, and $f_{r,\operatorname{wt}\left(r\right)}^{s2}$ are correctly defined as the frequency of unmutated genes in the library, and the counts $n_{r,\operatorname{wt}\left(r\right)}^{\rm{pre}}$, etc. are defined as the number of reads at $r$ attributable to unmutated genes. In this case, measurement of these counts requires sequencing with linkage as in \cite{hiatt2010parallel,wu2014hiv,kitzman2015massively}. The proper analysis of libraries containing only unmutated and singly mutated clones sequenced without linkage is beyond the scope of this paper.

If we knew the frequencies $\mu_{r,x}$, $f_{r,x}$, $f_{r,x}^{s1}$, and $f_{r,x}^{s2}$, we could calculate parameters that reflect the effects of mutations. One parameter that characterizes the effect of mutating $r$ from $\operatorname{wt}\left(r\right)$ to $x$ for the experiment in Figure \ref{fig:DeepMutationalScan}A is the \emph{enrichment ratio}, which is the relative frequency of mutations to $x$ after selection versus before selection:
\begin{equation}
\label{eq:enrichment}
\phi_{r,x} = \frac{f_{r,x} / f_{r,\operatorname{wt}\left(r\right)}}{\mu_{r,x} / \mu_{r,\operatorname{wt}\left(r\right)}}.
\end{equation}
Beneficial mutations have $\phi_{r,x} > 1$, while deleterious ones have $\phi_{r,x} < 1$. A related parameter is the \emph{preference} $\pi_{r,x}$ of $r$ for $x$. At each site, the preferences are simply the enrichment ratios rescaled to sum to one:
\begin{equation}
\label{eq:preference}
\pi_{r,x} = \frac{\phi_{r,x}}{\sum_y \phi_{r,y}} = \frac{f_{r,x} / \mu_{r,x}}{\sum_y f_{r,y} / \mu_{r,y}},
\end{equation}
or equivalently
\begin{equation}
\label{eq:preference2}
f_{r,x} = \frac{\pi_{r,x} \times \mu_{r,x}}{\sum_y \pi_{r,y} \times \mu_{r,y}},
\end{equation}
where $y$ is summed over all character identities (all nucleotides, codons, or amino acids).
The preferences can be intuitively visualized (Figure \ref{fig:DeepMutationalScan}A) and interpreted as the equilibrium frequencies in substitution models for gene evolution~\cite{bloom2014experimentally,bloom2014informed} (after accounting for uneven mutational rates~\cite{yampolsky2005exchangeability,stoltzfus2009climbing}).

@@

\subsubsection*{Implementation.}
The program \texttt{dms\_inferprefs} in the \texttt{dms\_tools} package infers the preferences by using \texttt{pystan}~\cite{pystan-software:2014} to perform MCMC over the posterior defined by the product of the likelihoods and priors in Equations \ref{eq:pr_nrerrpre}, \ref{eq:pr_nrerrpost}, \ref{eq:pr_nrpre}, \ref{eq:pr_nrpost}, \ref{eq:pr_pir}, \ref{eq:pr_mur}, \ref{eq:pr_epsilonr}, and \ref{eq:pr_rhor}. The program runs four chains from different initial values, and checks for convergence by ensuring that the Gelman-Rubin statistic $\hat{R}$~\cite{gelman1992inference} is $<1.1$ and the effective sample size is $>100$; the number of MCMC iterations is increased until convergence is achieved. The program \texttt{dms\_logoplot} in the \texttt{dms\_tools} package visualizes the posterior mean preferences via an extension to \texttt{weblogo}~\cite{crooks2004}. The program \texttt{dms\_merge} can be used to average preferences inferred from different experimental replicates that have individually been analyzed by \texttt{dms\_inferprefs}, and the program \texttt{dms\_correlate} can be used to compute the correlations among inferences from different replicates.

@@

Figures \ref{fig:MelnikovPrefs} and \ref{fig:ThyagarajanPrefs} also illustrate use of \texttt{dms\_correlate} to assess the correlation between preferences inferred from different biological replicates~\cite{blainey2014points} of the experiment. The inclusion and analysis of such replicates is the only sure way to fully assess the sources of noise associated with deep mutational scanning.

\subsubsection*{Testing on simulated data.}
To test the accuracy of preference-inference by \texttt{dms\_tools}, I simulated deep mutational scanning counts using the preferences in Figure \ref{fig:MelnikovPrefs}, both with and without errors quantified by appropriate controls. Importantly, the error and mutation rates for these simulations were \emph{not} uniform across sites and characters, but were simulated to have a level of unevenness comparable to that observed in real experiments. I then used \texttt{dms\_tools} to infer preferences from the simulated data, and also made similar inferences using simple ratio estimation (Equation \ref{eq:enrichmentcounts}). Figure \ref{fig:PrefCorrelations} shows the inferred preferences versus the actual values used to simulate the data. For simulations with mutation counts (quantified by the product $N\overline{\mu}$ of the depth and average per-site mutation rate) $\sim 1000$ to 2000, the inferences are quite accurate. Inferences made by \texttt{dms\_tools} are always more accurate than those obtained by simply taking ratios of mutation counts.

\subsubsection*{Is the Bayesian inference worthwhile?}
The foregoing sections explain why the Bayesian inference of preferences implemented in \texttt{dms\_tools} is conceptually preferable to estimating mutational effects via direct ratio estimation using Equation \ref{eq:enrichmentcounts}. However, do the practical benefits of this Bayesian inference justify its increased complexity? The simulations in the previous section show that the Bayesian inference is more accurate, but in the absence of background errors (Figure \ref{fig:PrefCorrelations}A) the magnitude of the improvement becomes negligible once the mutation counts per site $N \overline{\mu}$ start to exceed $\sim10^3$. When there is a need to correct for background errors (Figure \ref{fig:PrefCorrelations}B), meaningful benefits of the Bayesian inference over enrichment ratios extend to somewhat higher sequencing depths. Overall, it appears that Bayesian inference will always perform as well or better than ratio estimation, but that the tangible benefit becomes negligible at high sequencing depth. In that case, the user will have to decide if the increased computational runtime and complexity of the Bayesian inference is worth a small improvement in accuracy. Simpler ratio estimation can be performed using the \texttt{--ratio\_estimation} option of \texttt{dms\_inferprefs} or using an alternative program such as \texttt{Enrich}~\cite{fowler2011enrich}. When applying ratio estimation to data where some mutations have low counts, it is important to include pseudocounts (denoted by $\mathcal{P}$ in Equation \ref{eq:enrichmentcounts}) as a form of regularization to avoid estimating excessively high or low preferences at sites with limited counts.

@@

\subsubsection*{Implementation.}
The program \texttt{dms\_inferdiffprefs} in the \texttt{dms\_tools} package infers the differential preferences by performing MCMC over the posterior defined by the product of the likelihoods and priors in Equations \ref{eq:pr_nrerr}, \ref{eq:pr_nrstart}, \ref{eq:pr_nrs1}, \ref{eq:pr_nrs2}, \ref{eq:pr_controlpir}, \ref{eq:pr_xir}, \ref{eq:pr_frstart}, and \ref{eq:pr_deltapi}. The MCMC is performed as described for the preferences, and characters can again be any of nucleotides, amino acids, or codons. The program \texttt{dms\_logoplot} visualizes the posterior mean differential preferences via an extension to \texttt{weblogo}~\cite{crooks2004}. In addition, \texttt{dms\_inferdiffprefs} creates text files that give the posterior probability that $\Delta\pi_{r,x} > 0$ or $< 0$. These posterior probabilities are \emph{not} corrected to account for the fact that multiple sites are typically being examined, although by default the inferences are made using the regularizing prior in Equation \ref{eq:pr_deltapi}.

\subsection*{Inferring differential preference with \texttt{dms\_tools}.}
To test the accuracy of differential preference inference by \texttt{dms\_tools}, I simulated an experiment like that in Figure \ref{fig:DeepMutationalScan}B with the starting counts based on Melnikov \textit{et al}'s actual deep mutational scanning data of a Tn5 transposon~\cite{melnikov2014comprehensive}. As shown by Figure \ref{fig:MelnikovSimulatedDiffPrefs}, \texttt{dms\_inferdiffprefs} accurately infers the differential preferences at typical experimental depths. The results are easily visualized with \texttt{dms\_logoplot}. To provide a second illustration of differential preferences, Figure \ref{fig:WuDiffPrefs} shows an analysis of the data obtained by Wu \textit{et al} when they performed an experiment like that in Figure \ref{fig:DeepMutationalScan}B on nucleotide mutants of the influenza NS gene in the presence or absence of interferon treatment.

@@

\begin{figure}[h!]
%\centerline{\includegraphics[width=6.5in]{../../examples/Melnikov_et_al_Tn5/infer_prefs_on_simulated_data/correlations.pdf}}
\cprotect\caption{\label{fig:PrefCorrelations}\csentence{Accuracy of preference inference on simulated data.} Deep mutational scanning counts were simulated using the preferences in Figure \ref{fig:MelnikovPrefs}A and realistic mutation and error rates that were uneven across sites and characters as in actual experiments. The simulations were done {\bf (A)} without or {\bf (B)} with sequencing errors quantified by control libraries. Plots show the correlation between the actual and inferred preferences as a function of the product of the sequencing depth $N$ and the average per-site mutation rate $\overline{\mu}$; real experiments typically have $N\overline{\mu} \sim 1000$ to 2000 depending on the sequencing depth and gene length. Preferences are inferred using the full algorithm in \texttt{dms\_tools} (top panels) or by simply calculating ratios of counts (bottom panels) using Equation \ref{eq:enrichmentcounts} and its logical extension to include errors, both with a pseudocount of one. The \texttt{dms\_tools} inferences are more accurate than the simple ratio estimation, with both methods converging to the actual values with increasing $N\overline{\mu}$. Given files with the mutation counts, the plots in this figure can be generated as \texttt{prefs\_corr.pdf} and \texttt{ratio\_corr.pdf} with commands such as:
\begin{lstlisting}
dms_inferprefs pre.txt post.txt inferred_prefs.txt --ncpus -1
dms_inferprefs pre.text post.text ratio_prefs.txt --ratio_estimation 1
dms_correlate actual_prefs.txt inferred_prefs.txt prefs_corr --name1 "actual" --name2 "inferred" --corr_on_plot --r2
dms_correlate actual_prefs.txt ratio_prefs.txt ratio_corr --name1 "actual" --name2 "inferred" --corr_on_plot --r2
\end{lstlisting}
}
\end{figure}
